---
title: "Stat 3302 final 2"
author: "Madison Bailey (bailey.1439)"
date: "4/17/2022"
output: html_document
---
### Libraries
```{r, message = FALSE}
library(tidyverse)
library(nnet)
msleep
```

### Change vore to factor variable by amt of meat consumed
```{r}
msleep <- msleep %>% filter(is.na(conservation) == FALSE) %>% 
  #domesticated vs others
  mutate(dom = ifelse(conservation == "domesticated", FALSE, TRUE))

msleep_factor <- as.double(factor(msleep$dom, levels = c(FALSE, TRUE)))
dom_factor <- as.double(factor(msleep$vore, levels = c("domesticated", "en", "lc", "nt", "vu")))
```

???
## sleep_total multinomial model
```{r}
msleep %>% ggplot(aes(x = conservation, y = sleep_total)) +
  geom_boxplot()

multinom_model <- multinom(dom_factor ~ sleep_total, 
                          data = filter(msleep, is.na(sleep_total) == FALSE), 
                          Hess = TRUE)
multinom_model
```


## Statistical stuff

for domesticated vs. everyone else
```{r}
covariance <- round(solve(multinom_model$Hessian), 3)
alpha <- 0.2098 + 0.02402 - 0.21875 #wgere were these taken from?
exp(alpha)

alpha_matrix <- matrix(c(0, -1, 0, -1, 0, -1), nrow = 1, ncol = 6)
se_alpha <- sqrt(alpha_matrix %*% covariance %*% t(alpha_matrix))

# confidence interval :
LB <- alpha - (1.96 * se_alpha)
UB <- alpha + (1.96 * se_alpha)
exp(LB)
exp(UB)

p_value(multinom_model)
```


# body weight

```{r}
msleep %>% ggplot(aes(x = conservation, y = bodywt)) +
  geom_boxplot()

msleep %>% ggplot(aes(x = conservation, y = ln(bodywt))) +
  geom_boxplot()

multinom_model <- multinom(dom_factor ~ ln(bodywt), 
                          data = filter(msleep, is.na(bodywt) == FALSE),
                          Hess = TRUE)
multinom_model
```

# relative brain wt

```{r}
msleep <- msleep %>% mutate(
  rel_brainwt = 100 * brainwt / bodywt
) %>% filter(is.na(rel_brainwt) == FALSE)

msleep %>% ggplot(aes(x = conservation, y = rel_brainwt)) +
  geom_boxplot()

multinom_model_dom <- multinom(msleep_factor ~ rel_brainwt,
                                data = msleep, Hess = TRUE)
multinom_model_dom
multinom_model_cons <- multinom(dom_factor ~ rel_brainwt,
                                data = msleep,
                                Hess = TRUE)
multinom_model_cons
```
# compute confidence interval for meat

```{r}
covariance <- round(solve(multinom_model_dom$Hessian), 3)
alpha <- 
exp(alpha)

alpha_matrix <- matrix(c(0, 1), nrow = 1, ncol = 2)
se_alpha <- sqrt(alpha_matrix %*% covariance %*% t(alpha_matrix))

# confidence interval :
LB <- alpha - (1.96 * se_alpha)
UB <- alpha + (1.96 * se_alpha)
exp(LB)
exp(UB)

p_value(multinom_model_dom)
```
The chances a mammal is an meat-eater over an herbivore increases 2.1148 times for each one-unit increase in the ratio in 10 x brain weight : body weight. A 95% confidence interval is : (0.9703, 4.609), barely not significant. 

# confidence interval for omnivore

```{r}
covariance <- round(solve(multinom_model_cons$Hessian), 3)
alpha <- -1.0699 - 1.0866 - 0.3079
exp(alpha)

alpha_matrix <- matrix(c(0, -1, 0, -1, 0, -1), nrow = 1, ncol = 6)
se_alpha <- sqrt(alpha_matrix %*% covariance %*% t(alpha_matrix))

# confidence interval :
LB <- alpha - (1.96 * se_alpha)
UB <- alpha + (1.96 * se_alpha)
exp(LB)
exp(UB)
msleep %>% ggplot(aes(x = conservation, y = rel_brainwt)) +
  geom_boxplot()

p_value(multinom_model_cons)
```

The chances a mammal is an any other animal over an omnivore increases 0.085 times for each one-unit increase in the ratio in 100 x brain weight : body weight. A 95% confidence interval is : (0.009805, 0.7378) -> conclude there is a significant relationship between being an omnivore and relative brain weight. Moreover, looking at p-values, we note that there is not a significant difference between insectivores and omnivores. 



# sleep_cycle

```{r}
multinom_model <- multinom(msleep_factor ~ sleep_cycle, 
                                   data = msleep, Hess = TRUE)
msleep %>% ggplot(aes(x = conservation, y = sleep_cycle)) +
  geom_boxplot()

multinom_model
```
